### 2025-05-06 src/compiler.lagda.tex

\documentclass{article}

\usepackage[active,tightpage]{preview}
\PreviewEnvironment{prev}
\setlength\PreviewBorder{0pt}

\usepackage{agda}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathabx}

\usepackage{unicode-math}
\setmathfont{XITS Math}

\usepackage{newunicodechar}
\newunicodechar{≤}{\ensuremath{\leq}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{∸}{\ensuremath{\dotdiv}}
\newunicodechar{∀}{\ensuremath{\forall}}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
\newunicodechar{ℤ}{\ensuremath{\mathbb{Z}}}
\newunicodechar{δ}{\ensuremath{\delta}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{⊢}{\ensuremath{\vdash}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
\newunicodechar{⟶}{\ensuremath{\longrightarrow}}
\newunicodechar{ₛ}{\ensuremath{\textsubscript{s}}}
\newunicodechar{ᵛ}{\ensuremath{\textsuperscript{v}}}
\newunicodechar{ʳ}{\ensuremath{\textsuperscript{r}}}
\newunicodechar{₁}{\ensuremath{\textsubscript{1}}}
\newunicodechar{₂}{\ensuremath{\textsubscript{2}}}
\newunicodechar{⟨}{\ensuremath{\langle}}
\newunicodechar{⟩}{\ensuremath{\rangle}}
\newunicodechar{–}{\ensuremath{-}}

\begin{document}
\begin{prev}
\begin{code}
module compiler where

open import source
open import target
open import lib

infixr 1 _⇒ₛ_ 
infixl 2 _×_

-- Product and projection function
data _×_ (A B : Set) : Set where
    _,_ : A → B → A × B

π₁ : ∀ {A B} → A × B → A
π₁ (a , _) = a

π₂ : ∀ {A B} → A × B → B
π₂ (_ , b) = b

--  Type Interpretation
Compl : SD → Set
Compl sd = I sd

_×ₛ_ : (SD → Set) → (SD → Set) → SD → Set
(P ×ₛ Q) sd = P sd × Q sd

_⇒ₛ_ : (SD → Set) → (SD → Set) → SD → Set
(P ⇒ₛ Q) sd = ∀{sd'} → (sd ≤ₛ sd') → P sd' → Q sd'

Intcompl : SD → Set
Intcompl = R ⇒ₛ Compl

⟦_⟧ty : Type → SD → Set
⟦ comm ⟧ty = Compl ⇒ₛ Compl
⟦ intexp ⟧ty = Intcompl ⇒ₛ Compl
⟦ intacc ⟧ty = Compl ⇒ₛ Intcompl
⟦ intvar ⟧ty = ⟦ intexp ⟧ty ×ₛ ⟦ intacc ⟧ty
⟦ θ₁ ⇒ θ₂ ⟧ty = ⟦ θ₁ ⟧ty ⇒ₛ ⟦ θ₂ ⟧ty

-- Unit type for empty context
data ∅ : Set where
    unit : ∅

-- Context Interpretation
⟦_⟧ctx : Context → SD → Set
⟦ · ⟧ctx _ = ∅
⟦ Γ , A ⟧ctx sd = ⟦ Γ ⟧ctx sd × ⟦ A ⟧ty sd

⟦_⟧var : ∀ {Γ A sd} → A ∈ Γ → ⟦ Γ ⟧ctx sd → ⟦ A ⟧ty sd
⟦ Zero ⟧var (_ , a) = a
⟦ Suc B∈Γ ⟧var (γ , _) =  ⟦ B∈Γ ⟧var γ

⟦_⟧sub : ∀ {A B sd} → A ≤: B → ⟦ A ⟧ty sd → ⟦ B ⟧ty sd
⟦ ≤:-refl ⟧sub a = a
⟦ ≤:-trans A≤:A' A'≤:A'' ⟧sub a = ⟦ A'≤:A'' ⟧sub (⟦ A≤:A' ⟧sub a)
⟦ ≤:-fn A≤:A' B'≤:B ⟧sub a = λ sd≤ₛsd' a' → ⟦ B'≤:B ⟧sub (a sd≤ₛsd' (⟦ A≤:A' ⟧sub a'))
⟦ var-≤:-exp ⟧sub (exp , acc) = exp
⟦ var-≤:-acc ⟧sub (exp , acc) = acc

-- Functorality
fmap-⇒ : ∀ {P Q sd sd'} → (P ⇒ₛ Q) sd → sd ≤ₛ sd' → (P ⇒ₛ Q) sd'
fmap-⇒ θ p p' x = θ (≤ₛ-trans p p') x

fmap-Compl : ∀ {sd sd'} → Compl sd → sd ≤ₛ sd' → Compl sd'
fmap-Compl {sd} c (<-f f<f') = popto sd (<-f f<f') c
fmap-Compl {⟨ f , d ⟩} {⟨ f , d' ⟩} c (≤-d d≤d') = 
    adjustdisp-dec ((d' - d) d≤d') (-→≤ d≤d') 
        (I-sub {n = (d' - d) d≤d'} (n-[n-m]≡m d≤d') c)
        -- (sub I (-ₛ≡ 
        --     {n = (d' - d) d≤d'} 
        --         (n-[n-m]≡m d≤d')) c)

fmap-L : ∀ {sd sd'} → L sd → sd ≤ₛ sd' → L sd'
fmap-L (l-var sdᵛ sdᵛ≤ₛsd) sd≤ₛsd' = l-var sdᵛ (≤ₛ-trans sdᵛ≤ₛsd sd≤ₛsd')
fmap-L (l-sbrs) _ = l-sbrs

fmap-S : ∀ {sd sd'} → S sd → sd ≤ₛ sd' → S sd'
fmap-S (s-l l) sd≤ₛsd' = s-l (fmap-L l sd≤ₛsd')
fmap-S (s-lit lit) _ = s-lit lit

fmap-A : ∀ {A sd sd'} → ⟦ A ⟧ty sd → sd ≤ₛ sd' → ⟦ A ⟧ty sd'
fmap-A {comm}  = fmap-⇒ {Compl} {Compl}
fmap-A {intexp} = fmap-⇒ {Intcompl} {Compl}
fmap-A {intacc} = fmap-⇒ {Compl} {Intcompl}
fmap-A {intvar} ( e , a ) sd≤ₛsd' = ( fmap-A {intexp} e sd≤ₛsd' , fmap-A {intacc} a sd≤ₛsd' )
fmap-A {A ⇒ B} = fmap-⇒ {⟦ A ⟧ty} {⟦ B ⟧ty}

fmap-Γ : ∀ {Γ sd sd'} → ⟦ Γ ⟧ctx sd → sd ≤ₛ sd' → ⟦ Γ ⟧ctx sd'
fmap-Γ {·} unit _ = unit
fmap-Γ {Γ , A} (γ , a) p = fmap-Γ γ p , fmap-A {A} a p

sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] : ∀ {sd sd'} → sd ≤ₛ sd' → (δ₁≤δ₂ : suc (SD.d sd) ≤ SD.d sd') → sd ≤ₛ ((sd' -ₛ ((SD.d sd' - (suc (SD.d sd))) δ₁≤δ₂)) (-→≤ δ₁≤δ₂))
sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] {⟨ f , _ ⟩} {⟨ f' , _ ⟩} (<-f f<f') _ = <-f f<f'
sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] {⟨ f , d ⟩} {⟨ f , d' ⟩} (≤-d d≤d') δ₁≤δ₂ = ≤-d (suc-d≤d'→d≤d'-[d'-[suc-d]] δ₁≤δ₂)

new-intvar : ∀ sd → ⟦ intvar ⟧ty sd
new-intvar sd = ( e , a )
    where
        sd' : SD
        e : ⟦ intexp ⟧ty sd
        e sd≤ₛsd' β = β ≤ₛ-refl (r-s (s-l (l-var sd sd≤ₛsd')))
        a : ⟦ intacc ⟧ty sd
        a sd≤ₛsd' κ (≤-d {d = d'} {d' = d''} d'≤d'') r = {!   !}
        -- a {sd' = sd'} sd≤ₛsd' κ (≤-d {d = d'} {d' = d''} d'≤d'') r = assign-dec ((d'' - d') d'≤d'') (-→≤ d'≤d'') 
        --     (l-var sd' (≤-d (m≡n,p≤n→p≤m (n-[n-m]≡m d'≤d'') ≤-refl))) r (I-sub {n = (d'' - d') d'≤d''} ((n-[n-m]≡m d'≤d'')) κ)

assign : (sd : SD) → (sd' : SD) → (S ⇒ₛ Compl) sd → sd ≤ₛ sd' → R sd' → I sd'
assign ⟨ f , d ⟩ ⟨ f' , d' ⟩ β sd≤ₛsd' r with (≤-compare {suc d} {d'})
... | leq δ₁≤δ₂ = assign-dec ((d' - (suc d)) δ₁≤δ₂) 
                    (-→≤ δ₁≤δ₂) (l-var ⟨ f , d ⟩ (sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] sd≤ₛsd' δ₁≤δ₂)) r 
                    (β ((sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] sd≤ₛsd' δ₁≤δ₂)) 
                        (s-l (l-var ⟨ f , d ⟩ ((sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] sd≤ₛsd' δ₁≤δ₂)))))

... | geq δ₂≤δ₁ = assign-inc (((suc d) - d') δ₂≤δ₁) 
                    (l-var ⟨ f , d ⟩ (≤ₛ-trans sd≤ₛsd' +ₛ→≤ₛ)) r 
                    (β ((≤ₛ-trans sd≤ₛsd' +ₛ→≤ₛ)) 
                        (s-l (l-var ⟨ f , d ⟩ ((≤ₛ-trans sd≤ₛsd' +ₛ→≤ₛ)))))

use-temp : ∀ {sd sd'} → (S ⇒ₛ Compl) sd → sd ≤ₛ sd' → R sd' → I sd'
use-temp β sd≤ₛsd' (r-s s) = β sd≤ₛsd' s
use-temp {sd} {sd'} β sd≤ₛsd' (r-unary uop s) = 
    assign sd sd' β sd≤ₛsd' (r-unary uop s)
use-temp {sd} {sd'} β sd≤ₛsd' (r-binary s₁ bop s₂) = 
    assign sd sd' β sd≤ₛsd' (r-binary s₁ bop s₂)

⟦_⟧ : ∀ {Γ A} → (e : Γ ⊢ A) → (sd : SD) → ⟦ Γ ⟧ctx sd → ⟦ A ⟧ty sd
⟦ Var A∈Γ ⟧ sd γ = ⟦ A∈Γ ⟧var γ
⟦ Sub Γ⊢A A≤:B ⟧ sd γ = ⟦ A≤:B ⟧sub (⟦ Γ⊢A ⟧ sd γ)
⟦ Lambda f ⟧ sd γ {sd' = sd'} sd≤ₛsd' a = ⟦ f ⟧ sd' (fmap-Γ γ sd≤ₛsd' , a) 
⟦ App f e ⟧ sd γ = ⟦ f ⟧ sd γ (≤-d ≤-refl) (⟦ e ⟧ sd γ)
⟦ Skip ⟧ sd γ sd≤ₛsd' κ = κ
⟦ Seq c₁ c₂ ⟧ sd γ sd≤ₛsd' κ = ⟦ c₁ ⟧ sd γ sd≤ₛsd' (⟦ c₂ ⟧ sd γ sd≤ₛsd' κ)
⟦ NewVar c ⟧ sd γ {sd' = sd'} sd≤ₛsd' κ = assign-inc 1 (l-var sd' (≤-d +→≤)) (r-s (s-lit (pos 0))) (⟦ c ⟧ (sd' +ₛ 1) (fmap-Γ (( fmap-Γ γ sd≤ₛsd' , new-intvar sd)) (+ₛ→≤ₛ {sd'} {1})) ≤ₛ-refl (adjustdisp-dec 1 +→≤ʳ (sub I (-ₛ≡ {d' = SD.d sd' + 1} {n = 1} (n+m-m≡n {m = 1})) κ)))
⟦ Assign a e ⟧ sd γ sd≤ₛsd' κ = ⟦ e ⟧ sd γ sd≤ₛsd' (⟦ a ⟧ sd γ sd≤ₛsd' κ)
⟦ Lit i ⟧ _ _ _ κ = κ ≤ₛ-refl (r-s (s-lit i))
⟦ Neg e ⟧ sd γ sd≤ₛsd' κ = ⟦ e ⟧ sd γ sd≤ₛsd' (use-temp λ sd≤ₛsd' s → κ sd≤ₛsd' (r-unary UNeg s))    
⟦ Plus e₁ e₂ ⟧ sd γ p κ = ⟦ e₁ ⟧ sd γ p (use-temp (λ p' s₁ → ⟦ e₂ ⟧ sd γ (≤ₛ-trans p p') (use-temp (λ p'' s₂ →  κ (≤ₛ-trans p' p'') (r-binary (fmap-S s₁ p'') BPlus s₂))))) 
\end{code}
\end{prev}
\end{document}