### 2025-05-06 src/fin/compiler.lagda.tex

\documentclass{article}

\usepackage[active,tightpage]{preview}
\PreviewEnvironment{prev}
\setlength\PreviewBorder{0pt}

\usepackage{agda}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathabx}

\usepackage{unicode-math}
\setmathfont{XITS Math}

\usepackage{newunicodechar}
\newunicodechar{≤}{\ensuremath{\leq}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{∸}{\ensuremath{\dotdiv}}
\newunicodechar{∀}{\ensuremath{\forall}}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
\newunicodechar{ℤ}{\ensuremath{\mathbb{Z}}}
\newunicodechar{δ}{\ensuremath{\delta}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{⊢}{\ensuremath{\vdash}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
\newunicodechar{⟶}{\ensuremath{\longrightarrow}}
\newunicodechar{ₛ}{\ensuremath{\textsubscript{s}}}
\newunicodechar{ᵛ}{\ensuremath{\textsuperscript{v}}}
\newunicodechar{ʳ}{\ensuremath{\textsuperscript{r}}}
\newunicodechar{₁}{\ensuremath{\textsubscript{1}}}
\newunicodechar{₂}{\ensuremath{\textsubscript{2}}}
\newunicodechar{⟨}{\ensuremath{\langle}}
\newunicodechar{⟩}{\ensuremath{\rangle}}
\newunicodechar{–}{\ensuremath{-}}

\begin{document}
\begin{prev}
\begin{code}
module compiler where

open import source
open import target
open import lib

infixr 1 _⇒ₛ_ 
infixl 2 _×_

-- Product and projection function
data _×_ (A B : Set) : Set where
    _,_ : A → B → A × B

π₁ : ∀ {A B} → A × B → A
π₁ (a , _) = a

π₂ : ∀ {A B} → A × B → B
π₂ (_ , b) = b

--  Type Interpretation
Compl : SD → Set
Compl sd = I sd

_×ₛ_ : (SD → Set) → (SD → Set) → SD → Set
(P ×ₛ Q) sd = P sd × Q sd

_⇒ₛ_ : (SD → Set) → (SD → Set) → SD → Set
(P ⇒ₛ Q) sd = ∀{sd'} → (sd ≤ₛ sd') → P sd' → Q sd'

Intcompl : SD → Set
Intcompl = R ⇒ₛ Compl

⟦_⟧ty : Type → SD → Set
⟦ comm ⟧ty = Compl ⇒ₛ Compl
⟦ intexp ⟧ty = Intcompl ⇒ₛ Compl
⟦ intacc ⟧ty = Compl ⇒ₛ Intcompl
⟦ intvar ⟧ty = ⟦ intexp ⟧ty ×ₛ ⟦ intacc ⟧ty
⟦ θ₁ ⇒ θ₂ ⟧ty = ⟦ θ₁ ⟧ty ⇒ₛ ⟦ θ₂ ⟧ty

-- Unit type for empty context
data ∅ : Set where
    unit : ∅

-- Context Interpretation
⟦_⟧ctx : Context → SD → Set
⟦ · ⟧ctx _ = ∅
⟦ Γ , A ⟧ctx sd = ⟦ Γ ⟧ctx sd × ⟦ A ⟧ty sd

get-var : ∀ {Γ A sd} → A ∈ Γ → ⟦ Γ ⟧ctx sd → ⟦ A ⟧ty sd
get-var Zero     (_ , a) = a
get-var (Suc x) (γ , _) = get-var x γ

-- Functorality
fmap-⇒ : ∀ {P Q sd sd'} → (P ⇒ₛ Q) sd → sd ≤ₛ sd' → (P ⇒ₛ Q) sd'
fmap-⇒ θ p p' x = θ (≤ₛ-trans p p') x

fmap-Compl : ∀ {sd sd'} → Compl sd → sd ≤ₛ sd' → Compl sd'
fmap-Compl {sd} c (<-f f<f') = popto sd (<-f f<f') c
fmap-Compl {⟨ f , d ⟩} {⟨ f , d' ⟩} c (≤-d d≤d') = 
    adjustdisp-dec (≤→Fin (-→≤ {d'} {≤→Fin d≤d'}) ) (sub I ((-ₛ≡ {n = ≤→Fin (-→≤ {d'} {≤→Fin d≤d'})} (n-[n-m]≡m d≤d'))) c)

fmap-L : ∀ {sd sd'} → L sd → sd ≤ₛ sd' → L sd'
fmap-L (l-var sdᵛ sdᵛ≤ₛsd) sd≤ₛsd' = l-var sdᵛ (≤ₛ-trans sdᵛ≤ₛsd sd≤ₛsd') 
fmap-L l-sbrs _ = l-sbrs

fmap-S : ∀ {sd sd'} → S sd → sd ≤ₛ sd' → S sd'
fmap-S (s-l l) p = s-l (fmap-L l p)
fmap-S (s-lit lit) _ = s-lit lit

fmap-A : ∀ {A sd sd'} → ⟦ A ⟧ty sd → sd ≤ₛ sd' → ⟦ A ⟧ty sd'
fmap-A {comm}  = fmap-⇒ {Compl} {Compl}
fmap-A {intexp} = fmap-⇒ {Intcompl} {Compl}
fmap-A {intacc} = fmap-⇒ {Compl} {Intcompl}
fmap-A {intvar} ( e , a ) p = ( fmap-A {intexp} e p , fmap-A {intacc} a p )
fmap-A {A ⇒ B} = fmap-⇒ {⟦ A ⟧ty} {⟦ B ⟧ty}

fmap-Γ : ∀ {Γ sd sd'} → ⟦ Γ ⟧ctx sd → sd ≤ₛ sd' → ⟦ Γ ⟧ctx sd'
fmap-Γ {·} unit _ = unit
fmap-Γ {Γ , A} (γ , a) p = fmap-Γ γ p , fmap-A {A} a p

-- assign : (δ₁ : ℕ) → (δ₂ : ℕ) → (sd : SD) → (S ⇒ₛ Compl) sd → (sd' : SD) → sd ≤ₛ sd' → (δ₂ ≤ SD.d sd') → R sd' → I sd'
-- assign δ₁ δ₂ sd β sd' sd≤ₛsd' δ₂≤d' r with (≤-compare {δ₁} {δ₂})
-- ... | leq δ₁≤δ₂ = assign-dec (≤→Fin (≤-trans (-→≤ {δ₂} {≤→Fin δ₁≤δ₂}) δ₂≤d')) {!   !} r (β {!   !} {!   !})
-- ... | geq δ₂≤δ₁ = assign-inc (δ₁ - (≤→Fin δ₂≤δ₁)) (l-var sd (≤ₛ-trans sd≤ₛsd' +ₛ→≤ₛ)) r (β (≤ₛ-trans sd≤ₛsd' +ₛ→≤ₛ) ((s-l (l-var sd (≤ₛ-trans sd≤ₛsd' +ₛ→≤ₛ)))))

sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] : ∀ {sd sd'} → sd ≤ₛ sd' → (δ₁≤δ₂ : suc (SD.d sd) ≤ SD.d sd') → sd ≤ₛ (sd' -ₛ ≤→Fin (-→≤ {SD.d sd'} {≤→Fin δ₁≤δ₂})) 
sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] {⟨ f , d ⟩} {⟨ f' , d' ⟩} (<-f f<f') _ = <-f f<f'
sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] {⟨ f , d ⟩} {⟨ f , d' ⟩} (≤-d d≤d') δ₁≤δ₂ = ≤-d (suc-d≤d'→d≤d'-[d'-[suc-d]] δ₁≤δ₂)

assign : (sd : SD) → (sd' : SD) → (S ⇒ₛ Compl) sd → sd ≤ₛ sd' → R sd' → I sd'
assign ⟨ f , d ⟩ ⟨ f' , d' ⟩ β sd≤ₛsd' r with (≤-compare {suc d} {d'})
... | leq δ₁≤δ₂ = assign-dec (≤→Fin (-→≤ {d'} {≤→Fin δ₁≤δ₂})) (l-var ⟨ f , d ⟩ (sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] sd≤ₛsd' δ₁≤δ₂)) r (β (sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] sd≤ₛsd' δ₁≤δ₂) (s-l (l-var ⟨ f , d ⟩ (sd≤ₛsd'→sd≤ₛsd'-ₛ[d'-[suc-d]] sd≤ₛsd' δ₁≤δ₂))))
... | geq δ₂≤δ₁ = assign-inc ((suc d) - (≤→Fin δ₂≤δ₁)) (l-var ⟨ f , d ⟩ (≤ₛ-trans sd≤ₛsd' +ₛ→≤ₛ)) r (β (≤ₛ-trans sd≤ₛsd' +ₛ→≤ₛ) (s-l (l-var ⟨ f , d ⟩ (≤ₛ-trans sd≤ₛsd' +ₛ→≤ₛ))))

-- doubted

-- use-temp : ∀ {sd sd'} → (β : (S ⇒ₛ Compl) sd) → sd ≤ₛ sd' → (r : R sd') → I sd'
-- use-temp β sd≤ₛsd' (r-s s) = β sd≤ₛsd' s
-- use-temp {sd} {sd'} β sd≤ₛsd' (r-unary uop s) = assign (SD.d sd + 1) (SD.d sd') sd β sd' sd≤ₛsd' ≤-refl (r-unary uop s)
-- use-temp {sd} {sd'} β sd≤ₛsd' (r-binary s₁ bop s₂) = assign (SD.d sd + 1) (SD.d sd') sd β sd' sd≤ₛsd' ≤-refl (r-binary s₁ bop s₂)

use-temp : ∀ {sd sd'} → (S ⇒ₛ Compl) sd → sd ≤ₛ sd' → R sd' → I sd'
use-temp β sd≤ₛsd' (r-s s) = β sd≤ₛsd' s
use-temp {sd} {sd'} β sd≤ₛsd' (r-unary uop s) = assign sd sd' β sd≤ₛsd' (r-unary uop s)
use-temp {sd} {sd'} β sd≤ₛsd' (r-binary s₁ bop s₂) = assign sd sd' β sd≤ₛsd' (r-binary s₁ bop s₂)

⟦_⟧ : ∀ {Γ A} → (e : Γ ⊢ A) → (sd : SD) → ⟦ Γ ⟧ctx sd → ⟦ A ⟧ty sd
⟦ Var x ⟧ _ γ = get-var x γ
⟦ Lambda f ⟧ sd γ {sd' = sd'} p a = ⟦ f ⟧ sd' (fmap-Γ γ p , a) 
⟦ App f e ⟧ sd γ = ⟦ f ⟧ sd γ (≤-d ≤-refl) (⟦ e ⟧ sd γ)
⟦ Skip ⟧ _ _ _ γ = γ
⟦ Seq c₁ c₂ ⟧ sd γ sd' p = ⟦ c₁ ⟧ sd γ sd' (⟦ c₂ ⟧ sd γ sd' p)
⟦ Lit i ⟧ _ _ _ κ = κ ≤ₛ-refl (r-s (s-lit i))
⟦ Neg e ⟧ sd γ p κ = ⟦ e ⟧ sd γ p (use-temp λ p s → κ p (r-unary UNeg s))    
⟦ Plus e₁ e₂ ⟧ sd γ p κ = ⟦ e₁ ⟧ sd γ p (use-temp (λ p' s₁ → ⟦ e₂ ⟧ sd γ (≤ₛ-trans p p') (use-temp (λ p'' s₂ →  κ (≤ₛ-trans p' p'') (r-binary (fmap-S s₁ p'') BPlus s₂)))))   
\end{code}
\end{prev}
\end{document}